### 2022-04-19 Markdown/Manuscript/Child_RMDs/All_Content.Rmd
initial blank template

---
title: "All_Content"
author: "Author"
date: "`r Sys.Date()`"
output: html_document
---

```{r child="../Child_RMDs/MS_Body.Rmd", echo=FALSE}
```

\newpage
# References {-}

<div id="refs"></div>
```{r echo=FALSE}
if(!params$X@tables_in_body_lgl){
  child_tabs_1L_chr <- c("../Child_RMDs/MS_Cndl_Tabs_Not_In_Body.Rmd")
}else{
  child_tabs_1L_chr <- NULL
}
if(!params$X@figures_in_body_lgl){
  child_figs_1L_chr <- c("../Child_RMDs/MS_Cndl_Figs_Not_In_Body.Rmd")
}else{
  child_figs_1L_chr <- NULL
}
```

```{r, child = child_tabs_1L_chr, echo=FALSE}
```
```{r, child = child_figs_1L_chr, echo=FALSE}
```
\newpage
\appendix
\counterwithin{figure}{section}
\counterwithin{table}{section}

# Appendix

```{r child="../Child_RMDs/TA_Body.Rmd", echo=FALSE}
```

<!-- NOTE TO SELF: INCORPORATE TA_CODE.RMD -->