### 2022-04-19 Markdown/Manuscript/Child_RMDs/MS_Tables.Rmd

---
title: "MA_Results"
author: "Author"
date: "`r Sys.Date()`"
output: html_document
---

```{r child="XX_Funs.Rmd", echo=FALSE}
```

```{r echo=FALSE}
if(params$X@tables_in_body_lgl){
  child_docs_ls <- list(one = "../Child_RMDs/MS_Tables_1.RMD",
                      two = "../Child_RMDs/MS_Tables_2.RMD")
}else{
    child_docs_ls <- list(one = NULL,
                      two = NULL)
}
```