### 2022-04-19 Markdown/Manuscript/Parent_HTML/Parent_HTML.Rmd

---
bibliography: ../Data/references.bib
output: 
  bookdown::html_document2: 
    pandoc_args: 
      - "--metadata-file=../Header/header_common.yaml"
      - '--lua-filter=../Filters/scholarly-metadata.lua'
      - '--lua-filter=../Filters/author-info-blocks.lua'
params:
  X: NULL
abstract: "`r paste0(stringr::str_replace_all(stringr::str_replace_all(stringr::str_replace_all(stringr::str_replace_all(readLines('../Header/abstract.txt'),'\\\\\\\\textbf',''),'\\{','**'),'\\}','**'),'\\\\\\\\newline','\\\\n'),collapse=' ')`"
graphics: yes
csl: plos.csl
---

```{r echo = F}
library(ready4show)
```

```{r child="../Child_RMDs/All_Content.Rmd", echo=FALSE}
```