### 2022-04-19 Markdown/Manuscript/Parent_PDF/Main_Bookdown_PDF.Rmd

---
bibliography: ../Data/references.bib
output: 
  bookdown::pdf_document2: 
    includes:
      in_header: preamble.tex
    pandoc_args: 
      - "--metadata-file=../Header/header_common.yaml"
      - '--lua-filter=../Filters/scholarly-metadata.lua'
      - '--lua-filter=../Filters/author-info-blocks.lua'
    toc: no
params:
  X: NULL
abstract: "`r paste0(readLines('../Header/abstract.txt'),collapse=' ')`"
graphics: yes
csl: plos.csl
---

```{r echo = F}
## NOTE: EDITED TEMPLATE FILE To Add rotating and multirow packages as per:
# https://stackoverflow.com/questions/31503418/how-to-add-a-latex-package-to-r-journal-template-from-rticle-package-in-r
library(ready4show)
options(tinytex.verbose = TRUE)
library(xtable)
options(xtable.floating = FALSE)
options(xtable.timestamp = "")
```

```{r child="../Child_RMDs/All_Content.Rmd", echo=FALSE}
```