### 2022-04-20 Markdown/Manuscript/Child_RMDs/Doodles.Rmd
scribbles

---
title: "Doodles"
author: "Matthew Hamilton"
date: '2022-04-19'
output: html_document
---

ISPOR T & V: 
a multiapplication model, transparency, validation,
For a multiapplication model, transparency, validation,
and reporting are ongoing processes. The multiapplica-
tion model is described (transparency)24 and validated,25
and the descriptions and validations are continuously
updated as science and the model evolve.26 In addition,
each instantiation of the model is described, validated,
and reported as each application is done.27