### 2022-05-17 Markdown/Manuscript/Child_RMDs/Doodles.Rmd
intro done

---
title: "Doodles"
author: "Matthew Hamilton"
date: '2022-04-19'
output: html_document
---

10 rules [@Erdemir2020]
.....DARTH

[@Sampson2019] - 
development of networks of health economists that are working on a particular disease area. 
f reference models, open-source modelling, or model registries are to be part of the model development process, then standards need to be developed to guide analysts in achieving best practice. The availability of guidance can help to create confidence in researchers to pursue transparency. Yet, selecting the best standards to use can also be challenging [56]. Wider uptake—and appropriate use—of standards, by both modellers and journals, should be encouraged, and research should evaluate the suitability of guidance relating to transparency.
However, transparency depends on the code being well structured and clearly commented and a user or reviewer is required to have knowledge of the language in order to validate, use or adapt the model. There is a steep learning curve for analysts unfamiliar with script-based models
There may also be barriers in the communication of models due to a lack of knowledge for end users. However, this can be substantially mitigated with the use of open access interface packages, like the R package Shiny, which can be used to turn an R model into an interactive web application [58]. This allows a user without any R knowledge to use or, in part, validate the model.
Collaboration
----
@Erdemir2020] - 
We recommend the use of existing repositories for disseminating code and models.

there is a need for guidance, frameworks and templates
Guidance needed for implementation. Current transparency guidance not good enough.

ome resistence to adoption centers on.... [@Padula2017] . 
HE commercial...[@Feenstra2022],
   author]

A major shift in how health economic models are implemented.....They may also have educational value in developing the health systems modelling field [@Dunlop2017]. However, open source computational models also present a range of conceptual, technical, legal and resourcing issues compared to more traditional approaches to model development.

The health economics field i

cooperation [@Arnold2010]
coding framework / guidance [@Alarid2019]
10 rules [@Erdemir2020]
Sys review??

Maybe [@Feenstra2022] - 
Model complexity increases validation needs as does multi-purpose modelling
conceptual model, input data, computational model, operational model
International collaboration in validation
Need for documentation
Commercial ownership of HE models reduce scope for open-source
finances, accountability, and ownership barriers to OS
still not very widespread in HE decision model
would require more resources, could greatly increase transparency
distinguishing grades of open source
Model validation should hence become more structured and standardized, as well as more transparent.

Modular validation - Ghabri, Justifies extensive validation
Citing Dunlop - HE lagging behind other modelling fields

Australian
poor validation - ghabri

Over the last decade there have been
the potential to increase the transparency, reproducibility and reuse of health economic analyses.

Open source frameworks, which have been recommended for the development of mental health modelling field [@RN73], have the potential to help address a number of these challenges. Open source approaches can promote common standards and interoperability, improve model transparency and facilitate routine updates to rectify errors, update data and add features.

- Our project - readyforwhatsnext : open source model of the systems shaping the mental health of young people
- Lack of guidance on open source health economic models and in particular on recommended frameworks for implementing the computational model.
- Focus computational model. Most in need of novel guidance. Also: Of the three components that comprise a mental health computer simulation model - the conceptual model, the mathematical model and the computational model - only the first two are typically developed and reported in a manner that facilitates reuse, refinement and extension by others. This is problematic as the code implementation is where the theory represented by a simulation model is most fully elaborated.

ISPOR T & V: 
a multiapplication model, transparency, validation,
For a multiapplication model, transparency, validation,
and reporting are ongoing processes. The multiapplica-
tion model is described (transparency)24 and validated,25
and the descriptions and validations are continuously
updated as science and the model evolve.26 In addition,
each instantiation of the model is described, validated,
and reported as each application is done.27

CLIPPINGS
It is frequently claimed that better approaches to the organising and financing of mental health service systems can generate significant gains in population health (Saxena, 2007; Chisholm, 2016).  Over the last two decades, policymakers in Australia have acted on these claims by investing in a series of major systemic mental health reforms, with improving the mental health of young people a major focus (McGorry, 2017). The scope and complexity of these reforms has increased over time, with the Victorian Government’s current plan to overhaul mental health service delivery (Victoria, 2021) arguably the most ambitious to-date.

 Although Australian mental health policy and system design processes has a formally specified role for a robust model-based decision aid, this tool,  the National Mental Health Services Planning Framework, has limited capacity to meet many of the core decision-support requirements of systems planners (Commission, 2020). These requirements are diverse and extensive, but are likely to include models that can:

•	anticipate the distribution and evolution of risk factors and need for mental health care by region and population group;

•	identify the budget, equity and efficiency impacts of alternative resource allocations between mental health programs;

•	provide guidance about how services can be optimally configured to achieve outcomes such as minimising wait times, maximising health gain, meeting patient preferences and improving societal welfare; and

•	predict how the supply, location and behaviour of mental healthcare providers change in response to alternative incentives and conditions.

Although a number of these or similar types of models have been developed for the Australian context, the validity, utility and accessibility of these models vary considerably.

and to anticipate and mitigate risks to the successful implementation of these reforms

Developing theories that produce valid and useful predictions about the potential impacts and feasibility of phased, multi-component programs of policy and service system reform is most likely a major and long term challenge. This undertaking might become more tractable if it could be chunked into smaller tasks and progressed by a research community that, even when working independently, routinely test, re-use and extend each other's work.

However, developing  This leads to computational model issue. This leads to open source.

The absence of common standards and taxonomies make it hard to develop these more extensive models by simply combining existing mental health models.

 which can be used to generate credible predictions about how these systems might respond to phased, multi-component redesign strategies

Common standards and frameworks for how computational models are developed could address some of these challenges [@RN73].

Outputs from discrete projects should be reproducible, generalisable and able to be linked like Lego bricks to progressively assemble more complex and realistic representations of the systems being modelled.

 Similarly,

These knowledge gaps may partly explain why debates within the mental health policymaking community extend beyond normative issues (what should policy aim to achieve) to technical questions (which system financing and organising choices are most likely to achieve specified objectives). A more mature mental health systems modelling field might see the technical aspects of re-engineering mental health systems subject to the same stringent scientific standards as those required for the development of therapeutic goods. Yet,

If these issues can be successfully addressed, there are potential benefits to both decision makers (who will have access to more relevant and trustworthy model-based decision aids) and modellers (for whom complex models will become more tractable through improved access to relevant code and data).

Access to some model-based decision aids and their constituent algorithms and data is highly restricted. The relevance of models to current and future decision contexts can also be limited as mental health models are rarely routinely updated with new data and improved algorithms.

In addition to issues relating to improving the quality and usefulness of models that address narrowly defined topics, there are additional challenges concerning the development of models with much broader scope.

A number of high income countries have begun to increase public expenditures on mental health services[@RN26], and in some cases have embraced ambitious programs of major structural reforms to mental health service systems[REFSGOHERE]. However

In the limited cases where code is made publicly available it is typically in the form of programs (an algorithm that performs one task) rather than code libraries (where the primary algorithm is decomponsed into multiple component algorithms, that can be re-used for a potentially large number of other purposes).