### 2022-05-31 Markdown/Manuscript/Child_RMDs/MS_Available_Data_And_Resources.Rmd
circulated draft

---
title: "MS_Available_Data_And_Resources"
author: "Author"
date: "`r Sys.Date()`"
output: html_document
---

Development versions of all code repositories referenced in this article are available in https://github.com/ready4-dev/ . Archived code releases are available in https://zenodo.org/communities/ready4  .

All data repositories referenced in this article are available in https://dataverse.harvard.edu/dataverse/ready4 .