### 2022-05-31 Markdown/Manuscript/Child_RMDs/XX_Funs.Rmd

---
title: "XX_Funs"
author: "Author"
date: "`r Sys.Date()`"
output: html_document
---
```{r echo=FALSE}
library(magrittr)
# Adapted fom: https://seankross.com/2016/08/02/How-R-Packages-are-Licensed.html
sorting_hat <- function(student, houses){
  choice <- purrr::map_lgl(houses, grepl, x = student)
  if(!any(choice)){
    return("Other")
  } else {
    return((houses[choice])[1])
  }
}
get_license_share <- function(license_1L_chr = "GPL"){
  license_table <- table(utils::available.packages()[,"License"])
names(license_table) %<>%
  purrr::map_chr(function(x){
    sorting_hat(x, c("Apache", "Artistic", "CC", "BSD", "MIT", "GPL"))
  })  
license_tbl <- as.data.frame(license_table) %>%
  dplyr::rename(License = Var1) %>%
  dplyr::group_by(License) %>%
  dplyr::summarize(Freq = sum(Freq)) %>%
  dplyr::ungroup() %>%
  dplyr::arrange(dplyr::desc(Freq))
#licenses <- license_tbl$Freq
 share_1L_dbl <- round(license_tbl$Freq[license_tbl$License==license_1L_chr]/sum(license_tbl$Freq)*100,2) 
 return(share_1L_dbl)
}

```