### 2023-01-31 Markdown/Manuscript/Child_RMDs/MS_Tables_3.Rmd
section 2 of disc

---
title: "Untitled"
author: "Author"
date: "`r Sys.Date()`"
output: html_document
---

```{r checktb, tab.cap='Framework checklist applied to utility mapping study', tab.id = 'checktb', echo = F, results = 'asis'}
    #  a <- c("TEST1","TEST2")
    # b <- c("\\cite{ready4show2022}", "hjhjhj")
    # tmp <- data.frame(a,b)   
#     kableExtra::kbl(tmp, booktabs = T, escape = F) %>%
#  kableExtra::kable_styling(
#    latex_options = c("striped", "hold_position", "repeat_header")
#    ) %>%  
#   kableExtra::column_spec(2, width = "50em") %>%
# kableExtra::landscape()
checklist_df <- data.frame(Ref = c(paste0("A",1:7),#paste0("T",1:6),paste0("I",1:4),paste0("M",1:2),paste0("E",1:3),paste0("L",1:3), paste0("Y",1:2)
                                   paste0("R",1:9),
                                   paste0("U",1:4)),
                           Standard = c("Public",
                                        "Documentation",
                                        "Syntax",
                                        "Literate programming",
                                        "Code coverage",
                                        "Reproducibility",
                                        "Citation tools",#
                                        # Reuse
                                        "Code licenses",
                                        "Data terms",
                                        "Separation",
                                        "Encapsulated, inheriting",
                                        "Functions",
                                        "Libraries",
                                        "Test data",
                                        "Prediction tools",
                                        "User interface",
                                        ## Updatable
                                        "Version controlled",
                                        "Semantic versioning",
                                        "Continuously integrated",
                                        "Deprecation"
                                       ),
                           Assessment = c("Yes",
                                        "Yes",
                                        "Yes",
                                        "Yes",
                                        "No",
                                        "Yes",
                                        "Yes",
                                        # Reuse
                                        "Yes",
                                        "Yes",
                                        "Yes",
                                        "Yes", # "Yes",
                                        "Yes",
                                        "Yes",
                                        "Yes",
                                        "Yes",
                                        "No",
                                        ## Updatable
                                        "Yes",
                                        "Yes",
                                        "Yes",
                                        "Yes"
                                        ),
                           Description = c("Datasets are available at https://doi.org/10.7910/DVN/DKDIB0 and https://doi.org/10.7910/DVN/HJXYKQ. For details of where to find Development and Archive code see: https://www.ready4-dev.com/docs/getting-started/software/libraries/types/module/  https://www.ready4-dev.com/docs/getting-started/software/executables/programs/  https://www.ready4-dev.com/docs/getting-started/software/executables/subroutines/ ",
                                           "All code libraries have documenting websites with URLs that concatenate 'https://ready4-dev.github.io/' and the library name (e.g. https://ready4-dev.github.io/youthvars). All three Markdown programs are self-documenting, with one [@hamilton_matthew_2022_6129906] including additional instructions in a README file. Only one sub-routine [@matthew_p_hamilton_2022_5976988] is documented with a meaningful README file. All datasets have meaningful metadata descriptors.",
                                           "All libraries are authored with ready4pack [@ready4pack2022] to ensure a consistent house style. All libraries except [@matthew_p_hamilton_2021_5646669] use framework syntax, as does one program [@hamilton_matthew_2022_6129906].",
                                           "All programs use literate programming.",
                                           "No current reporting of code coverage.",
                                           "All parts of the study workflow from raw data ingest through to data processing, analysis, reporting and dissemination of study outputs can be reproduced (if granted access to source data) or replicated (using supplied synthetic data) with one program [@hamilton_matthew_2022_6129906].",
                                           "All libraries can return citation details when running R's 'citation' function. All code and data repositories have tools to generate citation details.",
                                           # Reuse
                                           "All code libraries, programs and sub-routines use GPL-3 licenses. ",
                                           "Datasets use amended version of template provided by Harvard Dataverse [@sampleterms2022].",
                                           "All development code is stored on repos in a GitHub organisation [@ready4gh2022] and all archived code releases are available in a Zenodo community [@ready4zen2022]. All non-confidential data are stored in repositories within a Harvard Dataverse collection [@ready4dv2022]. ",
                                           "Four [@hamilton_matthew_2022_6084467; @hamilton_matthew_2022_6084824; @hamilton_matthew_2022_6116701; @gao_caroline_2022_6130155] out of five libraries include encapsulating and inheriting modules. See: https://www.ready4-dev.com/docs/model/finding-modules/ ", 
                                           "All code libraries include functions. The most complete list of functions for each library is available by clicking the 'Manual - Developer (PDF)' link on each package's documentation homepage (see item A2 above). The relationship between functions can be illustrated using an app: https://www.ready4-dev.com/docs/getting-started/software/libraries/dependencies/",
                                           "All module data-structures and algorithms are distributed as five code libraries [@hamilton_matthew_2022_6084467; @hamilton_matthew_2022_6084824; @hamilton_matthew_2022_6116701; @gao_caroline_2022_6130155; @matthew_p_hamilton_2021_5646669].",
                                           "Two synthetic datasets and their data dictionaries are publicly available in a data repository [@DVN/HJXYKQ_2021]. One (ymh_clinical_tb.RDS) closely resembles the study dataset and was released so that the main study algorithm [@hamilton_matthew_2022_6129906] can be rerun by those without access to the confidential study dataset. The other (eq5d_ds_tb.Rds) is deliberately different to the source dataset in both variable naming convention and the concepts used for predictors and outcome measures. It was created to demonstrate the transferability of study algorithms.",                                           #
                                           "Model catalogues (PDF files beginning with 'AAA_TTU_MDL_CTG') are available in the study results dataset [@DVN/DKDIB0_2021] and describe the predictive performance of all models under a variety of usage regimes (including when the source dataset in the R model object is replaced with fake data). The youthu library [@matthew_p_hamilton_2021_5646669] includes tools for searching for and applying models compatible with different types of input data. An example program to demonstrate this functionality is available in both RMarkdown [@hamilton_matthew_2022_6416330] and rendered PDF formats (the 'Application.pdf' file in the study results dataset [@DVN/DKDIB0_2021]).  ",
                                           "No user interface has yet been developed.",
                                        ## Updatable
                                           "All code is version controlled using git [@git20XX] and GitHub [@github2007]. All source code is available in a GitHub organisation [@ready4gh2022].",
                                           "Semantic versioning is used in all code. As no code library has yet been submitted to CRAN [@CRAN2022], only the development version extensions of each version number have been incremented to date.",
                                           "All five libraries use continuous integration (CI). CI results for each library can be viewed at a URL that concatenates 'https://github.com/ready4-dev/', the library name and '/actions' (e.g. https://github.com/ready4-dev/youthvars/actions)",
                                           "Retired code is deprecated using tools from the lifecycle R library [@lifecycle2021] (e.g. everything after “## DEPRECATED FNS” in https://github.com/ready4-dev/youthvars/blob/main/data-raw/fns/add.R ). Package vignettes and datasets are also deprecated e.g. https://ready4-dev.github.io/youthvars/articles/Replication_DS.html )"
                                          ))

tb2_nms_chr<-colnames(checklist_df)
checklist_df <- data.frame(checklist_df[,1],
                           rep(ifelse(params$X@outp_formats_chr[1] == "Word"," ","&nbsp; "),
                               nrow(checklist_df)),
                           checklist_df[,2],
                           rep(ifelse(params$X@outp_formats_chr[1] == "Word"," ","&nbsp; &nbsp; &nbsp; &nbsp; "),
                               nrow(checklist_df)),
                           checklist_df[,3],
                           rep(ifelse(params$X@outp_formats_chr[1] == "Word"," ","&nbsp; "),
                               nrow(checklist_df)),
                 checklist_df[,4])
colnames(checklist_df) <- c(ifelse(params$X@outp_formats_chr[1] == "Word","Code"," "),
                            ifelse(params$X@outp_formats_chr[1] == "Word","YYY"," "),
                            "Standard",#
                             ifelse(params$X@outp_formats_chr[1] == "Word","ZZZ","   "),
                            "Met?",
                            ifelse(params$X@outp_formats_chr[1] == "Word","AAA","   "),
                            tb2_nms_chr[4])
if(params$X@outp_formats_chr[1] == "Word") {
  library("flextable")
  library("ftExtra")
  checklist_df[,
               c(1,3,5,7
                  )] %>%
    ready4show::print_table(output_type_1L_chr = params$X@outp_formats_chr[1],
                            caption_1L_chr = knitr::opts_current$get("tab.cap"),
                            mkdn_tbl_ref_1L_chr = paste0("tab:",knitr::opts_current$get("tab.id")),
                            use_rdocx_1L_lgl = ifelse(params$X@outp_formats_chr[1]=="Word",T,F),
                            add_to_row_ls = NULL,
                            sanitize_fn = NULL) %>%
  ftExtra::colformat_md()
  }else{
    checklist_df %>%
    kableExtra::kbl(format = "markdown",
                    booktabs = T,
                    caption = knitr::opts_current$get("tab.cap")) 
    }
```