### 2023-04-19 Markdown/Manuscript/Child_RMDs/MS_Conclusion.Rmd

---
title: "MS_Conclusion"
author: "Author"
date: "`r Sys.Date()`"
output: html_document
---
`r ifelse(is.na(params$X@conclusion_1L_chr),"",params$X@conclusion_1L_chr)` 
We have identified criteria that can be used to systematically assess extent to which the computational implementation of health economic models adheres to the ethical goals of transparency, reusability and updatability. We have developed an open-source software framework that can support the ethical computational implementation of economic models in youth mental health. Our framework can be used as a prototype for developing future software frameworks to support ethical implementation of CHEMs.