### 2023-08-01 Markdown/Manuscript/Child_RMDs/MS_Tables.Rmd
working my way through supervisors comments

---
title: "MA_Results"
author: "Author"
date: "`r Sys.Date()`"
output: html_document
---

```{r echo=FALSE}
if(params$X@tables_in_body_lgl){
  child_docs_ls <- list(one = "../Child_RMDs/MS_Tables_1.RMD",
                        two = "../Child_RMDs/MS_Tables_2.RMD",
                        three = "../Child_RMDs/MS_Tables_3.RMD",
                        four = "../Child_RMDs/MS_Tables_4.RMD")
}else{
    child_docs_ls <- list(one = NULL,
                      two = NULL,
                      three = NULL,
                      four = NULL)
}
```