### 2023-08-01 Markdown/Manuscript/Child_RMDs/MS_Tables_2.Rmd

---
title: "Predr_Cors"
author: "Author"
date: "`r Sys.Date()`"
output: html_document
---
```{r timelygls, tab.cap='Framework standards', tab.id = 'timelygls', echo = F, results = 'asis'}
library(kableExtra)  
df <- data.frame(`Objective` = rep("",20),
  # `Models are` = c(rep("Transparent",6),
  #                                 rep("Iterative",4),
  #                                 rep("Modular",2),
  #                                 rep("Epitomising",3),
  #                                 rep("Licensed",3),
  #                                 rep("Yielding",2)),
  #                Goal = c(rep("people can easily see how a model has been implemented and tested",6),
  #                         rep("a model is routinely updated to maintain and improve validity",4),
  #                         rep("models and their components can be combined to extend their scope",2),
  #                         rep("model code is sufficiently general to be be re-used in multiple decision contexts",3),
  #                         rep("a model, its components and derivatives are persistently re-usable by other modeller",3),
  #                         rep("a model can be simply, flexibly and reliably used as a decision aid",2)),
                 Standard = c(paste0("A",#"T", 
                                      1:7
                                      ),
                               paste0("R",1:9),# paste0("M",1:2),
                               # paste0("E",1:3),
                               # paste0("L",1:3),
                               # paste0("Y",1:2),
                               paste0("U",#"I",
                                      1:4)
                               ),
                 Meaning = c(
                   "Open online repositories are used to permanently archive, uniquely identify and transparently record authorship and development history of model code and data.",#T1
                   # T2 MISSING - Merged With T1
                   "Model code and data are documented.",#<T2, Was T3
                   "Model code uses consistent and abstracted syntax.",# T3, Was T4 -->
                   "Literate programming is used to implement model analyses.",#T4,Was T5
                   "Model code coverage is reported.", #T5, New, replaces version now moved to T4 
                   "All parts of a study analysis and reporting workflow can be reproduced and/or replicated.", #T6, was T7, Replaces Deleted T6 
                   "Model code and data are distributed with tools to support appropriate citation.",#*L3: New, new location
                   "Model code is made available for re-use under copyleft or permissive licenses.",#L1: **
                   "Non-confidential model data is licensed for liberal re-use (potentially subject to ethical use terms).",#L2
                   "Model code and data are stored and managed separately.",#*M1: Modified (lost ref to licensed and cited) 
                   "Model code uses encapsulation and inheritence for data structures.",#*M2: Replacement - Old version now merged with E1 # And with E2: "Modules are inheriting data-structures.",#E2
                   "Model code uses functions to implement algorithms.",#E1: Modified (lost ref to OOP and merged with M2) -->
                   "Model code is distributed as code libraries.",
                   "Test data is available to demonstrate the transferability of model code.",#E3: **
                   # OLD CITATION
                   "Statistical models are distributed with tools for making out of sample predictions.",#*Y1: New, Replaces Old Y1 (now Y2)
                   "User-interfaces allow non-technical users to configure and use models.",#Y2: Was Y1     
                   "Model code and data are version controlled.",#*I1, Was I2 
                   "The signficiance and status of code and data updates are indicated with semantic versioning and release types.",#I2, Was I1 
                   "Continuous integration is used to verify model code updates.",#I3 
                   "Deprecation conventions are used to retire model code and data."#**I4 
                 ),
                 check.names = F)
if(params$X@outp_formats_chr[1] == "Word") {
  library("flextable")
  library("ftExtra")
  df %>%
    ready4show::print_table(output_type_1L_chr = params$X@outp_formats_chr[1],
                            caption_1L_chr = knitr::opts_current$get("tab.cap"),
                            mkdn_tbl_ref_1L_chr = paste0("tab:",knitr::opts_current$get("tab.id")),
                            use_rdocx_1L_lgl = ifelse(params$X@outp_formats_chr[1]=="Word",T,F),
                            add_to_row_ls = NULL,
                            sanitize_fn = NULL) %>%
    ftExtra::colformat_md()
  }else{
    df %>%
      kableExtra::kbl(booktabs = T, 
                      escape = F,
                      caption = knitr::opts_current$get("tab.cap")) %>%
      kableExtra::kable_paper(full_width = F) %>%
      kableExtra::column_spec(1, bold=T) %>%
      # kableExtra::collapse_rows(columns = 1) %>%
      kableExtra::column_spec(3, width = "35em") %>%
      kableExtra::pack_rows("Accountable - it is easy to see who developed, tested and applied a model and how they did it.",1,7) %>%
      kableExtra::pack_rows("Updatable - a model and its components are maintained and continuously improved.",17,20) %>%
      # kableExtra::pack_rows("Modular - models and their components can be combined to extend their scope",11,12) %>%
      # kableExtra::pack_rows("Epitomising - model code can be re-used in other decision contexts",13,15) %>%
      kableExtra::pack_rows("Reusable - a model and its components can be used in other models and by other modellers.",8,16) #%>%
      # kableExtra::pack_rows("Yielding - a model can be simply, flexibly and reliably used as a decision aid",19,20)

    }
```