### 2023-08-14 Markdown/Manuscript/Parent_PDF/Test.Rmd

---
# Supported options: 
#   sn-nature:       Style for submissions to Nature Portfolio journals
#   sn-basic:        Basic Springer Nature Reference Style/Chemistry Reference Style
#   sn-mathphys:     Math and Physical Sciences Reference Style
#   sn-aps:          American Physical Society (APS) Reference Style
#   sn-vancouver:    Vancouver Reference Style
#   sn-apa:          APA Reference Style 
#   sn-chicago:      Chicago-based Humanities Reference Style
#   default:         Default

classoptions: 
  - sn-basic      
  # - Numbered      # Optional: Use numbered references instead of namedate references (only for sn-nature, sn-basic, sn-vancouver, sn-chicago, sn-mathphys or sn-nature)
  # - referee       # Optional: Use double line spacing 
  # - lineno        # Optional: Add line numbers
  # - iicol         # Optional: Double column layour

title: Article Title
titlerunning: Article Title runing

authors: 
  - prefix: Dr.
    firstname: Leading
    particle: van 
    lastname: Author
    suffix: III
    naturename: Poet Laureate
    degrees: MSc, PhD
    email: abc@def
    affiliation: [1,2]
    corresponding: TRUE
    equalcont: These authors contributed equally to this work.

  - firstname: Second
    lastname: Author
    affiliation: 2
    equalcont: These authors contributed equally to this work.

affiliations:
  - number: 1
    corresponding: TRUE
    info:
      orgdiv: Department
      orgname: Organization
    address:
        street: Street
        city: City
        postcode: 100190
        state: State
        country: Country
  - number: 2
    corresponding: TRUE
    info:
      orgname: Other Organisation

keywords:
  - key
  - dictionary
  - word

pacs: 
  jel:
    - "D8"
    - "H51"
  msc:
    - "35A01"
    - "65L10"

# Sample for unstructured abstract 
# abstract: |
#   The abstract serves both as a general introduction to the topic and as a brief, non-technical summary of the main results and their implications. Authors are advised to check the author instructions for the journal they are submitting to for word limits and if structural elements like subheadings, citations, or equations are permitted.

# Sample for structured abstract
abstract: |
  **Purpose**: The abstract serves both as a general introduction to the topic and as a brief, non-technical summary of the main results and their implications. The abstract must not include subheadings (unless expressly permitted in the journal's Instructions to Authors), equations or citations. As a guide the abstract should not exceed 200 words. Most journals do not set a hard limit however authors are advised to check the author instructions for the journal they are submitting to.

  **Methods:** The abstract serves both as a general introduction to the topic and as a brief, non-technical summary of the main results and their implications. The abstract must not include subheadings (unless expressly permitted in the journal's Instructions to Authors), equations or citations. As a guide the abstract should not exceed 200 words. Most journals do not set a hard limit however authors are advised to check the author instructions for the journal they are submitting to.

  **Results:** The abstract serves both as a general introduction to the topic and as a brief, non-technical summary of the main results and their implications. The abstract must not include subheadings (unless expressly permitted in the journal's Instructions to Authors), equations or citations. As a guide the abstract should not exceed 200 words. Most journals do not set a hard limit however authors are advised to check the author instructions for the journal they are submitting to.

  **Conclusion:** The abstract serves both as a general introduction to the topic and as a brief, non-technical summary of the main results and their implications. The abstract must not include subheadings (unless expressly permitted in the journal's Instructions to Authors), equations or citations. As a guide the abstract should not exceed 200 words. Most journals do not set a hard limit however authors are advised to check the author instructions for the journal they are submitting to.}

bibliography: ../Data/references.bib

# header-includes: |
#   %% Per the spinger doc, new theorem styles can be included using built in style, 
#   %% but it seems the don't work so commented below
#   %\theoremstyle{thmstyleone}%
#   \newtheorem{theorem}{Theorem}%  meant for continuous numbers
#   %%\newtheorem{theorem}{Theorem}[section]% meant for sectionwise numbers
#   %% optional argument [theorem] produces theorem numbering sequence instead of independent numbers for Proposition
#   \newtheorem{proposition}[theorem]{Proposition}%
#   %%\newtheorem{proposition}{Proposition}% to get separate numbers for theorem and proposition etc.
# 
#   %% \theoremstyle{thmstyletwo}%
#   \theoremstyle{remark}
#   \newtheorem{example}{Example}%
#   \newtheorem{remark}{Remark}%
# 
#   %% \theoremstyle{thmstylethree}%
#   \theoremstyle{definition}
#   \newtheorem{definition}{Definition}%
params:
  X: NULL
# output: 
#   bookdown::pdf_book: 
#     base_format: rticles::springer_article # rticles::arxiv_article
#     includes:
#       in_header: preamble.tex
output: rticles::springer_article
header-includes:
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage[normalem]{ulem}
- \usepackage{threeparttablex}
- \usepackage{makecell}
---

```{r echo = F}
## NOTE: EDITED TEMPLATE FILE To Add rotating and multirow packages as per:
# https://stackoverflow.com/questions/31503418/how-to-add-a-latex-package-to-r-journal-template-from-rticle-package-in-r
library(ready4show)
options(tinytex.verbose = TRUE)
library(xtable)
options(xtable.floating = FALSE)
options(xtable.timestamp = "")
```
TEST