### 2023-10-08 Markdown/Manuscript/Child_RMDs/MS_Tables_4.Rmd

---
title: "Untitled"
author: "Author"
date: "`r Sys.Date()`"
output: html_document
---

```{r checktb, tab.cap='Transparent, reusable and updatable (TRU) assessment criteria applied to outcome valuation computational health economic model (CHEM)', tab.id = 'checktb', echo = F, results = 'asis'}
library(kableExtra)  
df <- data.frame(FIRST = c("REPLACE", "THESE", "PLACEHOLDERS"),
                 SECOND = c("A","B","C"),
                 THIRD = c("ONE","TWO","THREE"),
                 check.names = F)
if(params$X@outp_formats_chr[1] == "Word") {
  library("flextable")
  library("ftExtra")
  df %>%
    ready4show::print_table(output_type_1L_chr = params$X@outp_formats_chr[1],
                            caption_1L_chr = knitr::opts_current$get("tab.cap"),
                            mkdn_tbl_ref_1L_chr = paste0("tab:",knitr::opts_current$get("tab.id")),
                            use_rdocx_1L_lgl = ifelse(params$X@outp_formats_chr[1]=="Word",T,F),
                            add_to_row_ls = NULL,
                            sanitize_fn = NULL) %>%
    ftExtra::colformat_md()
  }else{
    df %>%
      kableExtra::kbl(booktabs = T, 
                      escape = F,
                      caption = knitr::opts_current$get("tab.cap")) %>%
      kableExtra::kable_paper(full_width = F) %>%
      kableExtra::column_spec(1, bold=T) %>%
      kableExtra::column_spec(3, width = "35em")
    }
```