### 2023-10-21 Markdown/Manuscript/Child_RMDs/MS_Aims.Rmd
one table to go

---
title: "MS_Aims"
author: "Author"
date: "`r Sys.Date()`"
output: html_document
---

In this paper, we:

(i)	propose a set of ethical responsibilities for CHEM project teams and criteria for assessing ethical CHEM implementations;

(ii) describe a prototype software framework we have developed to support the ethical implementation of CHEMs; and

(iii)	assess an initial set of youth mental health CHEMs developed with our software framework against our proposed ethical CHEM implementation criteria.