### 2023-10-21 Markdown/Manuscript/Child_RMDs/MS_Tables_4.Rmd

---
title: "Untitled"
author: "Author"
date: "`r Sys.Date()`"
output: html_document
---

```{r checktb, tab.cap='Assessment of a CHEM implementation in youth mental health using Transparent, Reusable and Updatable (TRU) criteria', tab.id = 'checktb', echo = F, results = 'asis'}
library(kableExtra)  
df <- data.frame(Criterion = c("T1 Software files are open access", 
                               "T2 Judgments and contributions are easily identified", 
                               "R1 Programming practices promote generalisability and transferability",
                               "R2 Licenses restrict only unethical reuse",
                               "U1 Maintenance infrastructure is in place",
                               "U2 New releases are systematically retested and deprecated"),
                 `Met?` = c(rep("Yes",5),"Partial"),
                 Detail = c("All source code and testing procedures are available in public GitHub repositories, with each code release persistently available in a Zenodo repository. Study dataset contains confidential patient records and was not published. Instead, a synthetic representation of the study dataset is persistently available in a repository in the Harvard Dataverse. Data files to support out of sample application of models are published at the same location.",
                            "All module libraries, programs, and datasets are distributed with citation information. GitHub repositories detail author code contributions over project development history are publicly visible. Model catalogues persistently available on the Harvard Dataverse describe the predictive performance of models under multiple usage regimes. Each code library is documented with worked examples of how to apply modules.  Analysis and reporting programs are self-documenting.  Sub-routines for generating reports are documented with README files.",
                            "Model code is written using both functional and object-oriented paradigms. Code library websites include hypothetical examples of generalisability (applying study algorithm to estimate mapping models from new data with the same predictor and outcome variables) and transferability (adapting study algorithm to develop mapping models from datasets with different predictor variables and outcomes measured with a different utility instrument)." ,
                            "All code is distributed using GPL-3 licenses. Datasets use amended version of template provided by Harvard Dataverse, allowing reuse of data subject to some ethical restrictions (e.g., use in efforts to reidentify study participants is prohibited).",
                            "All code is version controlled using git and GitHub, with semantic versioning. Each code library has a specified maintainer and guidance for potential code contributors is available on the project website.",
                            "Continuous integration used for all code libraries, primarily for acceptance testing. Only limited use is made of unit testing. Retired library code is deprecated using tools from the lifecycle R library. Library documentation articles and datasets are also deprecated."),
                 check.names = F)
if(params$X@outp_formats_chr[1] == "Word") {
  library("flextable")
  library("ftExtra")
  df %>%
    ready4show::print_table(output_type_1L_chr = params$X@outp_formats_chr[1],
                            caption_1L_chr = knitr::opts_current$get("tab.cap"),
                            mkdn_tbl_ref_1L_chr = paste0("tab:",knitr::opts_current$get("tab.id")),
                            use_rdocx_1L_lgl = ifelse(params$X@outp_formats_chr[1]=="Word",T,F),
                            add_to_row_ls = NULL,
                            sanitize_fn = NULL) %>%
    ftExtra::colformat_md()
  }else{
    df %>%
      kableExtra::kbl(booktabs = T, 
                      escape = F,
                      caption = knitr::opts_current$get("tab.cap")) %>%
      kableExtra::kable_paper(full_width = F) %>%
      kableExtra::column_spec(1, width = "10em", bold=T) %>%
      kableExtra::column_spec(3, width = "35em")
    }
```