### 2023-10-21 Markdown/Manuscript/Parent_PDF/Parent_PDF.Rmd

---
bibliography: ../Data/references.bib
output: 
  bookdown::pdf_book: 
    base_format: rticles::springer_article # rticles::arxiv_article
    includes:
      in_header: preamble.tex
    keep_tex: true
    pandoc_args: 
      - "--metadata-file=../Header/header_common.yaml"
      - '--lua-filter=../Filters/scholarly-metadata.lua'
      - '--lua-filter=../Filters/author-info-blocks.lua'
  # bookdown::pdf_document2: 
  #   includes:
  #     in_header: preamble.tex
  #   pandoc_args: 
  #     - "--metadata-file=../Header/header_common.yaml"
  #     - '--lua-filter=../Filters/scholarly-metadata.lua'
  #     - '--lua-filter=../Filters/author-info-blocks.lua'
  #   toc: no
  #   keep_tex: true

params:
  X: NULL
abstract: "TO BE REPLACED"
# abstract: "`r paste0(readLines('../Header/abstract.txt'),collapse=' ')`"
graphics: yes
#csl: plos.csl
title: REPLACE ME
titlerunning: TEMPORARY
authorrunning: Hamilton, Gao, Wiesner et al.
thanks: | 
    The study was funded by ADD DETAILS HERE. 
keywords: one, two, three, four, five, six
#PACS: 
#- PAC1
#- superPAC

# MSC:
# - MSC code 1
# - MSC code 2
classoptions: 
  - sn-basic      
  # - Numbered      # Optional: Use numbered references instead of namedate references (only for sn-nature, sn-basic, sn-vancouver, sn-chicago, sn-mathphys or sn-nature)
  # - referee       # Optional: Use double line spacing 
  # - lineno        # Optional: Add line numbers
  # - iicol         # Optional: Double column layour
#biblio-style: spbasic
# bibstyle options spbasic(default), spphys, spmpsci
# header-includes:
# - \usepackage{booktabs}
# - \usepackage{longtable}
# - \usepackage{array}
# - \usepackage{multirow}
# - \usepackage{wrapfig}
# - \usepackage{float}
# - \usepackage{colortbl}
# - \usepackage{pdflscape}
# - \usepackage{tabu}
# - \usepackage{threeparttable}
# - \usepackage[normalem]{ulem}
# - \usepackage{threeparttablex}
# - \usepackage{makecell}
---

```{r echo = F}
## NOTE: EDITED TEMPLATE FILE To Add rotating and multirow packages as per:
# https://stackoverflow.com/questions/31503418/how-to-add-a-latex-package-to-r-journal-template-from-rticle-package-in-r
library(ready4show)
options(tinytex.verbose = TRUE)
library(xtable)
options(xtable.floating = FALSE)
options(xtable.timestamp = "")
```

```{r child="../Child_RMDs/All_Content.Rmd", echo=FALSE}
```