### 2023-10-22 Markdown/Manuscript/Parent_PDF/Springer.Rmd

---
# Supported options: 
#   sn-nature:       Style for submissions to Nature Portfolio journals
#   sn-basic:        Basic Springer Nature Reference Style/Chemistry Reference Style
#   sn-mathphys:     Math and Physical Sciences Reference Style
#   sn-aps:          American Physical Society (APS) Reference Style
#   sn-vancouver:    Vancouver Reference Style
#   sn-apa:          APA Reference Style 
#   sn-chicago:      Chicago-based Humanities Reference Style
#   default:         Default

classoptions: 
  - sn-vancouver      
  - Numbered      # Optional: Use numbered references instead of namedate references (only for sn-nature, sn-basic, sn-vancouver, sn-chicago, sn-mathphys or sn-nature)
  # - referee       # Optional: Use double line spacing 
  # - lineno        # Optional: Add line numbers
  # - iicol         # Optional: Double column layour

title: A prototype software framework for transparent, reusable and updatable computational health economic models
titlerunning: Transparent, reusable and updatable computational models

authors: 
  - prefix: Dr.
    firstname: Leading
    particle: van 
    lastname: Author
    suffix: III
    naturename: Poet Laureate
    degrees: MSc, PhD
    email: abc@def
    affiliation: [1,2]
    corresponding: TRUE
    equalcont: These authors contributed equally to this work.

  - firstname: Second
    lastname: Author
    affiliation: 2
    equalcont: These authors contributed equally to this work.

affiliations:
  - number: 1
    corresponding: TRUE
    info:
      orgdiv: Department
      orgname: Organization
    address:
        street: Street
        city: City
        postcode: 100190
        state: State
        country: Country
  - number: 2
    corresponding: TRUE
    info:
      orgname: Other Organisation

keywords:
  - key
  - dictionary
  - word

pacs: 
  jel:
    - "D8"
    - "H51"
  msc:
    - "35A01"
    - "65L10"

# Sample for unstructured abstract

params:
  X: NULL

abstract: "`r paste0(readLines('../Header/abstract.txt'),collapse=' ')`"

bibliography: bibliography.bib

header-includes: |
  %% Per the spinger doc, new theorem styles can be included using built in style, 
  %% but it seems the don't work so commented below
  %\theoremstyle{thmstyleone}%
  \newtheorem{theorem}{Theorem}%  meant for continuous numbers
  %%\newtheorem{theorem}{Theorem}[section]% meant for sectionwise numbers
  %% optional argument [theorem] produces theorem numbering sequence instead of independent numbers for Proposition
  \newtheorem{proposition}[theorem]{Proposition}%
  %%\newtheorem{proposition}{Proposition}% to get separate numbers for theorem and proposition etc.

  %% \theoremstyle{thmstyletwo}%
  \theoremstyle{remark}
  \newtheorem{example}{Example}%
  \newtheorem{remark}{Remark}%

  %% \theoremstyle{thmstylethree}%
  \theoremstyle{definition}
  \newtheorem{definition}{Definition}%

  \usepackage{float}
  \let\origfigure\figure
  \let\endorigfigure\endfigure
  \renewenvironment{figure}[1][2] {
      \expandafter\origfigure\expandafter[H]
  } {
      \endorigfigure
  }
  \usepackage{lscape}
  \usepackage{pdflscape}
  \newcommand{\blandscape}{\begin{landscape}}
  \newcommand{\elandscape}{\end{landscape}}
  \usepackage[none]{hyphenat}
  \usepackage[export]{adjustbox}
  \usepackage{multirow}
  \usepackage{longtable}
  \usepackage{array}
  \usepackage{wrapfig}
  \usepackage{colortbl}
  \usepackage{booktabs} 
  \usepackage{colortbl}
  \usepackage{pdflscape}
  \usepackage{tabu}
  \usepackage{threeparttable}
  \usepackage[normalem]{ulem}
  \usepackage{threeparttablex}
  \usepackage{makecell}

# output: rticles::springer_article
output: 
  bookdown::pdf_book: 
    base_format: rticles::springer_article
    # includes:
    #   in_header: preamble.tex
    keep_tex: true
    pandoc_args: 
      - '--lua-filter=../Filters/scholarly-metadata.lua'
      - '--lua-filter=../Filters/author-info-blocks.lua'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      out.extra = "")
#library(flextable)# This forces knitr to label all figures.
```

```{r child="../Child_RMDs/All_Content.Rmd", echo=FALSE, warning=FALSE}
```