### 2024-01-24 Markdown/Manuscript/Child_RMDs/MS_Aims.Rmd
extracted ethics component

---
title: "MS_Aims"
author: "Author"
date: "`r Sys.Date()`"
output: html_document
---

In this paper, we:

(i)	describe a prototype software framework we have developed to support the ethical implementation of CHEMs; and
<!-- propose a set of ethical responsibilities for CHEM project teams and criteria for assessing ethical CHEM implementations; -->

<!-- (ii)  -->

(iii)	assess an initial set of youth mental health CHEMs developed with our software framework against our proposed ethical CHEM implementation criteria.