### 2024-01-24 Markdown/Manuscript/Child_RMDs/MS_Conclusion.Rmd

---
title: "MS_Conclusion"
author: "Author"
date: "`r Sys.Date()`"
output: html_document
---
`r ifelse(is.na(params$X@conclusion_1L_chr),"",params$X@conclusion_1L_chr)` 
<!-- We have identified criteria that can be used to systematically assess the extent to which the computational implementation of health economic models adheres to transparency, reusability and updatability criteria.  -->
We have developed a prototype software framework that can support the implementation of CHEMs that largely meet explicit criteria for CHEM transparency, reusability and updataility. Our framework can be used as a prototype for developing future software frameworks to support ethical implementation of CHEMs. Barriers to ethical CHEM practice arising from resource constraints should be addressed by research funders.