### 2024-02-16 Markdown/Manuscript/Child_RMDs/MS_Aims.Rmd
minor edits to body

---
title: "MS_Aims"
author: "Author"
date: "`r Sys.Date()`"
output: html_document
---

In this paper, we describe:

(i)	a prototype software framework we have developed to support the development of modular and open source CHEMs; and
<!-- propose a set of ethical responsibilities for CHEM project teams and criteria for assessing ethical CHEM implementations; -->
<!-- (ii)  -->
(ii)	how we have applied that software framework to develop and share an initial set of modules for a economic model of youth mental health.