### 2024-02-18 Markdown/Manuscript/Child_RMDs/MS_Results.Rmd
added motivation

---
title: "MA_Results"
author: "Author"
date: "`r Sys.Date()`"
output: html_document
---

# Motivation
```{r child="../Child_RMDs/MS_Imp_Motivation.Rmd", echo=FALSE}
```

# Prototype software framework
```{r child="../Child_RMDs/MS_Imp_Toolkit.Rmd", echo=FALSE}
```

# Application in youth mental health
```{r child="../Child_RMDs/MS_Imp_Example.Rmd", echo=FALSE}
```