### 2024-02-20 Markdown/Manuscript/Child_RMDs/MS_Imp_Toolkit.Rmd
minor

---
title: "MS_Imp_Guidelines"
output: html_document
date: '2022-07-21'
---

## Purpose
We decided to develop a software framework to help us address challenges we encountered during the planning and early implementation of readyforwhatsnext. The software framework is called ready4 [@ready424]. The purpose of ready4 is to support the development of CHEMs that meet explicit criteria for transparency, reusability and updatability. We have describe these criteria elsewhere [@ethicsCHEMS2024], but briefly they are: software files are open access, developer contributions are easily identified, programming practices promote selective reuse, licenses restrict only unethical reuse, maintenance infrastructure is in place and new releases are systematically retested, with deprecation of outdated artefacts.

## User requirements
We created ready4 with one primary user group in mind - ourselves. The user requirements (URs) for ready4 evolved over six years (2018-2024) as we became aware of new challenges in implementing readyforwhatsnext that we needed to solve. We specified these user requirements in terms that facilitate solutions that may meet the needs of two other CHEM developers. Currently, these requirements are:

UR 1) A *template* for creating CHEM modules that can: (i) safely share inputs and outputs with each other without impairing any module's integrity; (ii) be run as independent models; and (iii) be selectively replaced, updated or extended.

UR 2) *Developer tools* that integrate with online open source software development services to support streamlined workflows for authoring R libaries of CHEM modules that are: (i) created using the template module; (ii) written in a consistent house style; (iii) consistently documented; (iv) licensed for reuse; (v) bundled with citation data specifying authorship, funding source, version number and a Digital Object Identifier (DOI); and (vi) quality assured with publicly available acceptance and unit tests.

UR 3) A simple *syntax* that promotes support consistency in how algorithms associated with CHEM modules are described and used in health economic analysis programs.

<!-- UR 4) Integration with online open source software development services to streamline the *authoring, testing, documenting and sharing of CHEM modules*.  -->

UR 5) *Data management tools* that integrate with online data repository services to streamline the ingesting, labeling and sharing of publicly available and confidential data.

UR 6) *Reporting tools* that include: 
<!-- help generate tables for multiple outputs, retrieve templates, modify templates with authorship and results -->
(i) replicable health economic analysis programs; and (ii) reporting subroutines that.

UR 7) Tools to *find CHEM modules, datasets and replication programs*.

UR 8) Integration with tools for documenting open source software to help *develop and maintain a consolidated CHEM project website*.

## Implementation
<!-- The software framework aims to provide a toolkit for: (i) enabling modular implementation of CHEMs; (ii) authoring and documenting CHEM modules; (iii) managing the labelling and transfer of CHEM input and output data; and (iv) authoring reproducible analyses that apply CHEMs to compatible data. -->
We implemented the software framework as R [@RCORE2022] code libraries that integrate with a number of online services and which are supported by a documentation website.

### R libraries 
A library in the R language will typically depend on multiple other R libraries, many of which are likely to have different authors. As the number of third-party dependencies of an R library grows so does the fragility of that library (e.g., the library may cease to work as intended due to changes in one of its dependency libraries). To reduce the fragility of our framework we implemented it as multiple R libraries rather than one R library. In total we authored six novel R libraries to implement the ready4 framework, all of which have distinct purposes and dependencies (Table \@ref(tab:cpkgs)).

One library provides the foundation for the ready4 framework. <!-- In modular model implementations, modules need to be able to share inputs and outputs with each other and to be run as independent models [@barros2023empowering].  -->The ready4 foundation library defines a template CHEM module using R’s S4 class system, that supports an object-oriented programming paradigm. Two notable features of object-oriented programming are encapsulation [@8717448] and inheritence [@8717448]. Encapsulation allows a CHEM module to bundle model data and algorithms in a protected environment. Encapsulation can be used to help ensure that CHEM modules continue to work as intended when they are combined<!-- [@ready4oop2022] -->. Inheritence allows new CHEM modules to be created that inherit some or all of the properties of a parent module. Inheritence is useful in cases where it is necessary to update CHEM modules that are being combined to account for interaction effects. Inheritance also facilitates the selective editing of CHEM modules when transferring models to different decision contexts<!-- [@ready4oop2022] -->. The ready4 template CHEM module can therefore be used to create other CHEM modules that inherit common set of properties (meeting *UR 1*). One of these inherited properties is a novel syntax of 15 core commands that enable CHEM module algorithms to be consistently named (meeting *UR 3*). The foundation library also contains tools for retrieving web-based information on CHEM modules, datasets and analysis programs authored with ready4 (meeting *UR 7*) and tools for developing and maintaining a project documentation website (*UR 8*).

Three framework libraries are designed to meet *URs 2 and 4*. <!-- Define functional and object-oriented approaches, acceptance and unit testing. -->The R language supports functional and object-oriented programming paradigms [@10.1214/13-STS452]. Authoring with each of these paradigms is facilitated by a dedicated CHEM module authoring library. One module authoring library contains tools for writing functions in a consistent house style and automatically generating basic documentation for each function. A second module authoring library contains tools to help streamline and standardise the authoring of novel, documented CHEM modules from the template CHEM module. A third module authoring library provides tools for disseminating themed bundles of CHEM modules as R libraries that are:

-	documented (with a website and PDF manuals);

-	licensed (using the copyleft GNU GPL-3 [66] by default);

-	easily citable (citation information can be retrieved within an R session or from hosting repositories); and

-	quality assured (each update triggers continuous integration workflows, including any acceptance and unit tests created by module library authors).

A library for managing CHEM data contains tools for supplying CHEM modules with input data ingested from local (i.e., a user's computer) or remote (online repositories) locations, labelling CHEM module datasets and exporting CHEM module data to online repositories (meeting *UR 5*). A library for authoring reproducible analysis programs and reporting subroutines (*UR6*) contains tools to help write executables that apply CHEM modules to compatible datasets for the purpose of undertaking health economic analyses. The library supports the creation of analysis programs that are self-documenting (code is integrated with plain English explanations of what it does) and subroutines that can be used as templates that trigger the creation of explanatory documents (e.g., a scientific manuscript).

```{r, child = child_docs_ls$three, echo=FALSE}
```

###	Online services 
Framework libraries are designed to be used in conjunction with a number of third-party online services that users of the framework will need to establish and configure accounts with.

We created a GitHub organisation (a collection of code repositories) where source code that we author is stored and version controlled. We configured individual repositories in our GitHub organisation to implement continuous integration. By default, code libraries authored with our framework will use continuous integration to assess compliance with policies specified by the Comprehensive R Archive Network (CRAN) [@CRAN2022]. To track our code coverage, we linked our GitHub organisation to an account we established at codecov [@codecov_2022]. To facilitate the creation and hosting of documentation websites, we enabled GitHub Pages in each repository used for code library development.

We also created a Zenodo community - a collection of permanent, uniquely identified repositories. We then linked our Zenodo community and GitHub organisation so that every time we specify a version of code in one of our GitHub repositories as a “release”, a copy of that code is automatically created on Zenodo with a DOI. Finally, to manage model datasets, we created a dedicated collection within the Harvard Dataverse installation.

###	Documentation website
We developed a framework documentation website (https://www.ready4-dev.com) that provides guidance to model developers on how to use and contribute improvements to the ready4 software framework and CHEM modules developed with it. The documentation website is versioned, which means documentation relating to prior versions of framework software can be archived, retrieved and viewed.

The documentation website was developed using the Hugo framework [@hugo_2023], Docsy theme [@docsy_2023] and Algolia DocSearch [@algoliadocsearch_2023] and is hosted using the Netlify [@netlify_2023] service.  We used functions from our foundation framework library to partially automate website updates relating to available CHEM modules, datasets and analysis programs. We linked our Netlify account to our GitHub organisation so that the project website would automatically update whenever its  source code (publicly available in a GitHub repository) was edited.