### 2024-02-20 Markdown/Manuscript/Child_RMDs/MS_Aims.Rmd
moved URs

---
title: "MS_Aims"
author: "Author"
date: "`r Sys.Date()`"
output: html_document
---

In this paper, we describe:

(i)	our motivation for developing a transferable CHEM in youth mental health;

(ii) a prototype software framework we have developed to support the development of transferable CHEMs; and
<!-- propose a set of ethical responsibilities for CHEM project teams and criteria for assessing ethical CHEM implementations; -->
<!-- (ii)  -->
(ii)	how we have applied that software framework to develop and share an initial set of modules for our youth mental health CHEM.