### 2024-02-25 Markdown/Manuscript/Child_RMDs/MS_Imp_Example.Rmd
abstract updated

---
title: "MS_Imp_Guidelines"
output: html_document
date: '2022-07-21'
---

<!-- Although we are principally interested in using CHEMs to answer policy questions relating to the mental health of young people in Australia, we want to facilitate CHEM transferability to other jurisdictions.  -->
Our early use of the ready4 framework has focused on developing readyforwhatsnext modules for constructing, sharing and using utility mapping models in youth mental health.

## Model user requirements
The initial model user requirements (MURs) that we specified for the readyforwhatsnext modules we developed for utility mapping are:

MUR 1) *Variable validation tools* that: (i) define the allowable range (minimum and maximum) and numeric class (integer or double precision) associated with different psychological, functional and health utility summary measures appropriate for use in youth mental health samples; (ii) provide an informative error message that halts an analysis or reporting program's execution should impermissible values be assigned to each measure.

MUR 2) *Dataset description tools* that: (i) attach structured metadata to a human records dataset that specify the unique person identifier, group assignment and data collection round variables; and (ii) generate tabular and graphical summaries of dataset descriptive statistics.

MUR 3) *Multi-attribute instrument scoring tools* that: (i) attach structured metadata about an instrument (e.g. name, version, country, items, domains, parameter values for scoring algorithms) to a human records dataset; (ii) apply pre-written (for AQol-6D and EQ-5D instruments) or user-supplied instrument scoring functions to a human records dataset; and (iii) generate descriptive plots of a dataset's instrument scores for individual items, domains and totals (weighted and unweighted).

MUR 4) *Utility mapping model construction tools* that: (i) attach structured metadata about the target utility instrument and candidate predictor variables and covariates to be used in constructing utility mapping models to a human records dataset; (ii) generate descriptive tables, a correlation matrix and plots about the variables to be assessed for inclusion in models; (iii) attach structured metadata about the candidate utility mapping models to be assessed; (iv) generate tabular and graphical summaries of the performance of each combination of model, predictors and covariates specified for evaluation by either a user or by a default algorithm; (v) strip confidential records from the constructed models so that they are able to be publicly shared.

MUR 5) *Utility mapping study reporting tools* that: (i) generate from a reporting template a PDF catalogue summarising the predictive performance under multiple configurations of utility mapping models selected for public dissemination; (ii) generate from a reporting template (either a pre-written default or a user-supplied customisation) a scientific summary manuscript summarising a utility mapping study; (iii) share model catalogues and (if desired) scientific summary along with metadata about utility mapping models in an open access repository; and (iv) facilitate the authoring of a single consolidated study reproducibility program that implements all steps inclusive of data ingest, instrument scoring, model construction, report authoring and dissemination of study artefacts.

MUR 6) *Utility prediction tools* that: (i) search open access repositories for utility mapping models developed with readyforwhastnext; (ii) retrieve relevant metadata, including a link to the model catalogue for models specified by a user; (iii) applies selected utility mapping model to make out of sample predictions using a dataset provided by a user; and (iv) (where there is health utility data at two time points) convert predicted health utility scores to Quality Adjusted Life Years (QALYs).

## Model implementation

### Online service configuration
We established and configured accounts with the online services supported by ready4. We created a GitHub organisation (a collection of code repositories) where source code that we author is stored and version controlled at https://github.com/ready4-dev/. We configured individual repositories in our GitHub organisation to implement continuous integration to assess the compliance of our module libraries with policies specified by the Comprehensive R Archive Network (CRAN) [@CRAN2022]. <!-- To track our code coverage, we linked our GitHub organisation to an account we established at codecov [@codecov_2022].  --> To facilitate the creation and hosting of module library websites, we enabled GitHub Pages in each code repository. We also created a Zenodo community, a collection of permanent, uniquely identified repositories, at https://zenodo.org/communities/ready4/. We then linked our Zenodo community and GitHub organisation so that every time we specify a version of code in one of our GitHub repositories as a “release”, a copy of that code is automatically archived on Zenodo with a DOI. Finally, to manage model datasets, we created a dedicated collection (https://dataverse.harvard.edu/dataverse/ready4) within the Harvard Dataverse installation.

<!-- To track our code coverage, we linked our GitHub organisation to an account we established at codecov [@codecov_2022].  -->
<!-- The framework also supports integrations with the open science repositories Zenodo [@Zenodo2013] and Dataverse [@Dataverse2007] that provide persistent storage solutions that generate a Digital Object Identifier (DOI) for each code and data collection. -->
<!-- The third party services that our framework is designed to work with include GitHub, Zenodo, Dataverse, -->
<!-- codecov [@codecov_2022] -->

### Modules
We used all six ready4 framework libraries to author five development version readyforwhatsnext module libraries (Table \@ref(tab:rfwnlibs)). These libraries are youthvars [@hamilton_matthew_2022_6084467] which addresses MURs 1 and 2, scorz [@hamilton_matthew_2022_6084824] which addresses MUR 3, specific [@hamilton_matthew_2022_6116701] which addresses MUR 4, TTU [@gao_caroline_2022_6130155] which addresses MUR 5 and youthu [@matthew_p_hamilton_2021_5646669] which addresses MUR 6.

### Reporting templates
We used the framework reporting tools library to author two reporting templates that are designed to be used in conjunction with these utility mapping models. The first reporting template is used to create a catalogue of utility mapping models [@hamilton_matthew_2022_6116385]. The second reporting template can be used to generate a manuscript [@matthew_p_hamilton_2022_5976988] providing a scientific summary of each study implemented with these modules.

### Application
We have previously described a study [@Hamilton2021.07.07.21260129] that used applied these modules and reporting templates to develop utility mapping models in a sample of young people presenting to primary mental health services. We used these module libraries to write programs that reproduced that study's reporting and analysis algorithm [@hamilton_matthew_2022_6129906], to demonstrate how to applying the utility mapping models constructed in that study to new data [@hamilton_matthew_2022_6416330] and to generate a synthetic representation of the study dataset [@hamilton_matthew_p_2022_6321821]. We also created an open access study input parameters and results dataset [@DVN/DKDIB0_2021]) and toy datasets to help demonstrate the transferability of the study algorithm [@DVN/HJXYKQ_2021]).

```{r, child = child_docs_ls$threeb, echo=FALSE}
```

## Documentation website
We used using the search and documentation tools from our foundation framework library to
create and help maintain a documentation website (https://readyforwhatsnext.org/) that includes details on available model modules, datasets and analysis programs and tutorials on using model modules.

## Transferability assessment
We subjectively assessed these utility mapping model artefacts against <!-- Edit_start -->our TRU criteria <!-- Edit_end --> (Table \@ref(tab:checktb)). For each criterion, we provided a global assessment of whether it was met using the responses “yes”, “no” or “partial”.  We believe our utility mapping study modules have satisfactorily met four of the six criteria (T1, R1, R2 and U1) and have partially met two criteria (T2 and U2). The main shortcomings that we identified when applying the assessment criteria was that we have yet to: (i) ensure that module library documentation includes examples of how to apply all exported methods and functions; and (ii) adequately implement unit testing.

\blandscape

```{r, child = child_docs_ls$four, echo=FALSE}
```

\elandscape