### 2024-02-27 Markdown/Manuscript/Child_RMDs/MS_Available_Data_And_Resources.Rmd
cull

---
title: "MS_Available_Data_And_Resources"
author: "Author"
date: "`r Sys.Date()`"
output: html_document
---

The ready4 framework is documented at https://www.ready4-dev.com. Documentation on readyforwhatsnext is available at https://readyforwhatsnext.org/. Development versions of all code repositories referenced in this article are available in https://github.com/ready4-dev/. Archived code releases are available in https://zenodo.org/communities/ready4. All data repositories referenced in this article are available in https://dataverse.harvard.edu/dataverse/ready4.