### 2024-02-27 Markdown/Manuscript/Child_RMDs/MS_Conclusion.Rmd

---
title: "MS_Conclusion"
author: "Author"
date: "`r Sys.Date()`"
output: html_document
---
`r ifelse(is.na(params$X@conclusion_1L_chr),"",params$X@conclusion_1L_chr)` 
<!-- We have identified criteria that can be used to systematically assess the extent to which the computational implementation of health economic models adheres to transparency, reusability and updatability criteria.  -->
Our software framework provides the outline of a technical solution to some of the challenges of transferable CHEM development using open-source and modular approaches. Our framework can be used as a prototype for developing future software frameworks to support implementation of transferable CHEMs. 
<!-- Barriers to ethical CHEM practice arising from resource constraints should be addressed by research funders. -->