### 2024-03-01 Markdown/Manuscript/Child_RMDs/MS_Aims.Rmd
ready for comparison

---
title: "MS_Aims"
author: "Author"
date: "`r Sys.Date()`"
output: html_document
---

In this paper, we describe:

(i)	our motivation for developing a transferable CHEM in youth mental health;

(ii) a prototype software framework we have developed to support the development of transferable CHEMs; and
<!-- propose a set of ethical responsibilities for CHEM project teams and criteria for assessing ethical CHEM implementations; -->
<!-- (ii)  -->

(iii)	how we have applied the prototype software framework to undertake initial development of our youth mental health CHEM.